### 2013-11-30 paper/paper.tex
Added new resources for compiling final report. View README instructions inside the paper folder for how to build the paper.

%\documentclass[3p]{elsarticle}
\documentclass[5p]{elsarticle}
\usepackage{natbib,graphicx,amsmath, mhchem,url,colortbl,fancyhdr,subfigure}
\usepackage{multicol}

\journal{EECS 545}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\begin{document}
\begin{frontmatter}

%\input{header.tex}
\title{Methods for Polyphonic Music Transcription} % Remove Piano part?
\author[add]{Jeremy Nash}
\ead{nashj@umich.edu}
\author[add]{Mark Liu}
\ead{markmliu@umich.edu}
\author[add]{Paul Schroeder}
\ead{pschro@umich.edu}
\address[add]{Electrical Engineering Department, University of Michigan, Ann Arbor, MI 48109}

%\input{abstract.tex}
\begin{abstract}
This paper presents a review of modern methods for polyphonic music transcription. NNN
\end{abstract}
\begin{keyword}
Transcription \sep sparse coding \sep non-negative matrix factorization \sep Bayesian non-parametrics \sep music information retrieval
\end{keyword}

\end{frontmatter}

\setcounter{topnumber}{1}

\section{Motivation}
% Explain like you're at a whiteboard

% Here's why people want to do this
% Study of music - understanding music
% recommender systems
% Music information database
% For learning musicians

% Here's why it's difficult
% Musical timbres
% Time varying - attack, decay, sustain, release
% Number of sources?
% Human perception of pitch
% Musical instrument tuning
% Completely new instruments - software created

% Here's a teaser on how I'm going to solve this problem

% Here's what I'm about to show you

\section{Problem Statement}

\section{Related Work}
% Contain a table with metrics data, brief description, year published

\section{Methodologies}

The two main methodologies in modern polyphonic music transcription reflect the supervised/unsupervised learning dichotomy in machine learning. The first approach learns separate discriminative models for the presence of a note over spectral features.

% Constant Q transform

% Semitone filter bank

\subsection{Spectrogram}
% Hint at the matrix factorization approach
\subsection{Constant Q transform}
\subsection{Semitone filter bank}
\subsection{Discriminative models}
% SVM, logistic regression
\subsection{SVM}
\subsection{Feed forward neural network}

\subsection{Non-negative matrix factorization}
% Non-negative Matrix factorization
\subsection{Bayesian nonparametric models}
\subsection{Smoothing}
\subsubsection{HMM smoothing}
\subsubsection{Probabilistic spectral smoothness}
\subsection{Recurrent neural network}

\section{Evaluation}

\section{Conclusion}
% Which method is fastest

% Which method has the best performance

% Which one is the simplest method

% Future directions

\section{Individual Effort}
\begin{description}
\item[$\bullet$ Jeremy] wrote and evaluted the SVM method in \cite{poliner2006discriminative}, the Bayesian nonparametric method in \cite{blei2010bayesian}, and the sparse non-negative matrix factorization method in \cite{abdallah2004polyphonic}. Jeremy also wrote the paper.
\item[$\bullet$ Mark] implemented the hidden Markov model in \cite{poliner2006discriminative} and the LSTM network in \cite{bock2012polyphonic}.
\item[$\bullet$ Paul] implemented the LSTM network in \cite{bock2012polyphonic}.
\end{description}

%    Problem statement, both conceptual and mathematical.
%    Motivation: why is the problem important.
%    Related work.
%    Methodologies explored and developed, and their advantages and disadvantages
%    Evaluation: experiments on sythetic and real data, performance measures used, results. I hope to see quantitative, objective evaluations, and comparisons with competitors
%    Conclusions: what did you learn, what were the project's success and failures?
%    Description of individual effort: At the end of the report, in a paragraph (included in the page limit), please include a brief description of each project member's contribution to the project.

% Support vector machines + HMM

% Honglak Lee's approach: use knn or RBM for input

% Non-negative matrix factorization

% Supervised NMF

% Combination NMF + svm

% Recurrent neural network (LSTM)

% Bayesian non-parametric method

% Bayesian non-parametric II (infinite hidden markov model)

% Our approach: Bayesian non parametrics with stochastic ...

%\input{1.tex}
%\input{fig1.tex}
%\input{2.tex}
%\input{3.tex}
%\input{fig2.tex}
%\input{4.tex}
%\input{5.tex}

% Your project should involve experiments that compare a few different algorithms. These should be well known or state-of-the-art methods, plus possibly any new methods you develop. The experiments should use real (as opposed to simulated) data sets. There is a large repository of machine learning data sets called the ``UCI Machine Learning Repository.'' There are other more domain-specific sources as well that you might find by searching.

%Your project should focus on a topic that has previously been studied in the machine learning literature. I include this requirement because in the past, some groups have defined entirely new problems for themselves, and been unable to make any progress.
%Your project report should contain an extensive literature review that summarizes fundamental ideas and state-of-the-art methods, with critiques of their strengths and weaknesses. My favorite reports to read are the ones that teach me something interesting.
%You are also encouraged to develop your own methods to solve your chosen problem. 
%Your project should involve experiments that compare a few different algorithms. These should be well known or state-of-the-art methods, plus possibly any new methods you develop. The experiments should use real (as opposed to simulated) data sets. There is a large repository of machine learning data sets called the ``UCI Machine Learning Repository.'' There are other more domain-specific sources as well that you might find by searching.

%Every report should address the following (but don't feel obligated to make these the section headings)
%    Problem statement, both conceptual and mathematical.
%    Motivation: why is the problem important.
%    Related work.
%    Methodologies explored and developed, and their advantages and disadvantages
%    Evaluation: experiments on sythetic and real data, performance measures used, results. I hope to see quantitative, objective evaluations, and comparisons with competitors
%    Conclusions: what did you learn, what were the project's success and failures?
%    Description of individual effort: At the end of the report, in a paragraph (included in the page limit), please include a brief description of each project member's contribution to the project.

\bibliographystyle{model1-num-names}
\bibliography{transcription}
\end{document}